---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval = FALSE}
library(BNPGrangerTestR)
```

To explore the basic capabilities of `BNPGrangerTestR`, we'll use the dataset 
`example_dataset`. This is a simulated sample with 200 registers and 2 
continuous responses.

`BNPGrangerTestR` aims to provide tools for testing the existence of a 
Granger causal relationship between any pair of variables, 
using a flexible Bayesian model.

### Fitting the model

First, we need to train the model:
```{r, message = FALSE, eval = FALSE}
object <-
  example_dataset |>
  bnpgrangertest()
```

### Testing for Granger causal relationships

Once the model is trained, we can compute the posterior probability of 
each Granger causal relationship as follows:

```{r, eval = FALSE}
summarize_gamma(object)
```

In this row of this tibble, `cause_var` is the causing variable, `effect_var`
is the affected variable, and `prob` is the posterior probability of `cause_var` 
causing `effect_var`. In this case, there is strong evidence of a `y1` 
(the 1st variable) Granger-causing `y2` (the 2nd variable), 
but not the other way around.

### Computing the IRF

In a similar, we can summarize the posterior distribution 
of the IRFs as follows:

```{r, eval = FALSE}
summarize_irf(object)
```

In each row of this tibble, `horizon` is the IRF horizon, `cause_var` and
`effect_var` are defined as before, `irf` is the posterior mean of the IRF,
and `lb` and `ub` represent the 90% credible interval of the IRF.

### Computing the posterior predictive pdf

Finally, we can summarize all (1 dimensional) posterior 
predictive pdfs as follows:

```{r, eval = FALSE}
summarize_pdf(object)
```

In each row of this tibble, `horizon` is the prediction horizon, `var` is the
target variable, `y` is the point for which the pdf is evaluated, and `pdf`
is the value of the posterior predictive pdf at `y`.

### Plotting capabilities

Instead of summarizying the results as a `tibble`, you can summarize them 
as a `ggplot` using the following functions:

```{r, eval = FALSE}
plot_gamma(object)
plot_irf(object)
plot_pdf(object)
```

### Extracting the MCMC chains

Finally, you can extract the raw chains of the main variables of our model
using the following functions:

```{r, eval = FALSE}
extract_chain_gamma(object)
extract_chain_irf(object)
extract_chain_pdf(object)
```
